---
title: "KKT with Weights"
author: "Sahir"
date: "May 16, 2017"
output:
  html_document:
    toc: true
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
rm(list=ls())
```

## Overview

This document verifies the KKT conditions for `glmnet` without and with weights. Note that `glmnet` internally rescales the observation weights to sum to the number of predictors. Let $N_T$ be the total number of observations and $w_i$ the observation weights for subject $i$. The following code rescales the weights:

```{r, eval=FALSE}
wi_scaled <- wi / sum(wi) * n
```

Let $\widetilde{w}_i$ be the re-scaled weights. Then the KKT equation is given by:

$$
\frac{1}{\sum_i^{N_T} \widetilde{w}_i} \sum_i^{N_T} \widetilde{w}_i \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j
$$
When all the weights are equal, $\sum_i^{N_T} \widetilde{w}_i = N_T$, and this equation simplifies to

$$
\frac{1}{N_T}  \sum_i^{N_T} \frac{w_i}{\sum_i^{N_T} w_i} N_T  \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j\\
\frac{w_i}{\sum_i^{N_T} w_i}  \sum_i^{N_T}  \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j\\
\frac{w_i}{N_T \cdot w_i}  \sum_i^{N_T}  \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j\\
\frac{1}{N_T }  \sum_i^{N_T}  \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j
$$

$$
\frac{1}{N_T}  \sum_i^{N_T}  \widetilde{X}_{ij} \left( \widetilde{Y}_i - \sum_{j=0}^{p}\widetilde{X}_{ij+1}\beta_j \right) = \lambda \gamma_j
$$


## Simulate Data




```{r}
library(glmnet)
library(MASS)
library(magrittr)

set.seed(1234)
# intercept
b0 <- 3
# true betas
b <- c(runif(10, 0.8,1.2), rep(0,180), runif(10, -1.2, -0.8))
# number of predictors
p <- length(b)
# sample size
n <- 100
# independent predictors
X <- mvrnorm(n, rep(1,p), diag(p))
# response
Y <- b0 + X %*% b + rnorm(n)

# threshold
threshold <- 1e-20

# kkt tolerance
tol.kkt <- 1e-9
```


## KKT with Equal Weights

```{r}
# weights
wi <- rep(1, n)
# wi <- sample(1:10, size = n, replace = T)

(m2 <- glmnet(y = Y, x = X, standardize = F, thresh = threshold, weights = wi))
(m2 <- glmnet(y = Y, x = X, standardize = F, thresh = threshold))
max(m2$lambda)
B <- as.matrix(m2$beta)

out_print <- matrix(NA, nrow = length(m2$lambda), ncol = 3,
                    dimnames = list(colnames(B),
                                    c("kkt_beta_nonzero",
                                      "kkt_beta_subgr", 
                                      "sum_wi_scaled")))
```



```{r, echo=FALSE}
for (lambda in colnames(B)) {
  # lambda = colnames(B)[1]
  beta <- B[, lambda, drop = F]
  lambda_value <- m2$lambda[which(lambda==colnames(B))]
  n <- nrow(X)

  # scale the weights to sum to nobs
  wi_scaled <- as.vector(wi) / sum(as.vector(wi)) * n
  wi_mat <- diag(wi_scaled)

  # KKT for beta
  # g0 <- (1 / sum(wi_scaled)) * crossprod(X * wi_scaled, (Y - X %*% beta - drop(m2$a0[lambda])))
  g0 <- (1 / sum(wi)) * crossprod(X * wi, (Y - X %*% beta - drop(m2$a0[lambda])))
  
  g <- g0 - lambda_value * sign(beta)

  # this is for when beta=0 and should be between -1 and 1
  gg <- g0 / lambda_value

  # which of the betas are non-zero
  oo <- abs(beta) > 0

  # if all betas are 0 then set to TRUE, else abs(g[oo]) will give error since 'oo' is all FALSE
  out_print[lambda, "kkt_beta_nonzero"] <- if (all(!oo)) 0 else sum(abs(g[oo]) > tol.kkt)
  out_print[lambda, "kkt_beta_subgr"] <- sum(abs(gg[!oo]) > 1)
  out_print[lambda, "sum_wi_scaled"] <- sum(wi_scaled)
  
  if (sum(abs(g[oo]) > tol.kkt) > 0) plot(abs(g[oo]))

}

out_print

mysd <- function(y) sqrt(sum((y - mean(y)) ^ 2)/length(y))
sy <- as.vector(scale(Y, scale = mysd(Y)))
sx <- scale(X, scale = apply(X, 2, mysd))
sx <- as.matrix(sx, ncol = p, nrow = n)
dim(sx)
(lambda.max <- max(abs(colSums(wi * sx * drop(sy)))) / n)
m2$lambda

sx <- scale(X,scale=apply(X,2,mysd))
sx <- as.matrix(sx,ncol=p,nrow=n)
sy <- as.vector(scale(Y, scale = mysd(Y)))
max(abs(colSums(sx*sy)))/100
fitGLM <- glmnet(sx, sy, standardize = F)
max(fitGLM$lambda)


rm(list = ls())
set.seed(1)
library("glmnet")
x <- matrix(rnorm(100*20),100,20)
y <- rnorm(100)
fitGLM <- glmnet(x,y)
max(fitGLM$lambda)



```


## KKT with Different Weights

```{r}
# weights
wi <- sample(1:10, size = n, replace = T)

(m2 <- glmnet(y = Y, x = X, standardize = F, thresh = threshold, weights = wi))
B <- as.matrix(m2$beta)

out_print <- matrix(NA, nrow = length(m2$lambda), ncol = 3,
                    dimnames = list(colnames(B),
                                    c("kkt_beta_nonzero",
                                      "kkt_beta_subgr", "sum_wi_scaled")))
```


```{r}
for (lambda in colnames(B)) {
# lambda=colnames(B)[1]
  beta <- B[, lambda, drop = F]
  lambda_value <- m2$lambda[which(lambda==colnames(B))]
  n <- nrow(X)
  
  # scale the weights to sum to nobs
  wi_scaled <- as.vector(wi) / sum(as.vector(wi)) * n
  wi_mat <- diag(wi_scaled)

  # KKT for beta
  # g0 <- (1 / sum(wi_scaled)) * crossprod(X * wi_scaled, (Y - X %*% beta - drop(m2$a0[lambda])))
  g0 <- (1 / sum(wi)) * crossprod(X * wi, (Y - X %*% beta - drop(m2$a0[lambda])))

  g <- g0 - lambda_value * sign(beta)

  # this is for when beta=0 and should be between -1 and 1
  gg <- g0 / lambda_value

  # which of the betas are non-zero
  oo <- abs(beta) > 0

  # if all betas are 0 then set to TRUE, else abs(g[oo]) will give error since 'oo' is all FALSE
  out_print[lambda, "kkt_beta_nonzero"] <- if (all(!oo)) 0 else sum(abs(g[oo]) > tol.kkt)
  out_print[lambda, "kkt_beta_subgr"] <- sum(abs(gg[!oo]) > 1)
  out_print[lambda, "sum_wi_scaled"] <- sum(wi_scaled)
  
  if (sum(abs(g[oo]) > tol.kkt) > 0) plot(abs(g[oo]))

}

(lambda.max <- max(abs(colSums(((1 / sum(wi_scaled)) * (wi_scaled * X) * drop(Y))))))

(lambda.max <- max(abs(colSums(((1 / sum(wi)) * (wi * X) * drop(Y))))))
  
  
m2$lambda

out_print
```


